---
title: "Project 1 - Karen Lopez"
date: "September 18, 2020"
output:
  rmarkdown::github_document:
    toc: true
knit: (function(input, ...){
  rmarkdown::render(
    input,
    output_file="README.md",
    envir=globalenv()
  )})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Reading and Summarizing Data from the NHL's API
This vignette was compiled to provide detailed instructions on reading and summarizing data from the National Hockey League's, or NHL's, API using R code.

## Required Packages
First, install and read in the necessary packages, as shown below. Packages only need to be installed once, but code is included in comment.  
```{r packages, message=FALSE}
#install.packages("tidyverse")
library(tidyverse)
library(httr)
library(jsonlite)
library(dplyr)
library(knitr)
```

## Make Contact and Get Data
This code contains functions that will contact the [NHL records API](https://gitlab.com/dword4/nhlapi/-/blob/master/records-api.md) and [NHL stat API](https://gitlab.com/dword4/nhlapi/-/blob/master/stats-api.md), and return parsed, well-formatted data.  

### Franchise API
The user is required to input team ID when using the last 3 functions listed here.
```{r franchiseAPI, message=FALSE}
#function that creates the url for the records API
getUrl <- function(input){
  baseUrl <- "https://records.nhl.com/site/api/franchise"
  fullURL <- paste0(baseUrl, input) #create url
}

franchise <- function(){
  franchiseAPI <- GET(getUrl(NULL)) #get url & data
  franchiseAPI #check connection
  franchiseText <- content(franchiseAPI, "text") #convert to JSON text form
  franchiseList <- fromJSON(franchiseText, flatten=TRUE) #convert form
  franchiseDF <- as.data.frame(franchiseList) #convert to df, adds .data to col names
  colnames(franchiseDF) <- c("id", "firstSeasonId", "lastSeasonId", "mostRecentTeamId", "teamCommonName", "teamPlaceName", "total") #rename columns
  
  return(franchiseDF)
}
franchiseData <- franchise() #save data frame object (used in other functions later)
print(kable(head(franchiseData))) #view first 6 rows

teamTotals <- function(){
  teamAPI <- GET(getUrl("-team-totals")) #get url & data
  teamText <- content(teamAPI, "text") #convert to JSON text form
  teamList <- fromJSON(teamText, flatten=TRUE) #convert form
  teamDF <- as.data.frame(teamList) #convert to df, adds .data to col names
  colnames(teamDF) <- c("id", "activeFranchise", "firstSeasonId", "franchiseId", "gameTypeId", "gamesPlayed", "goalsAgainst", "goalsFor", "homeLosses", "homeOvertimeLosses", "homeTies", "homeWins", "lastSeasonId", "losses","overtimeLosses", "penaltyMinutes", "pointPctg", "points", "roadLosses", "roadOvertimeLosses", "roadTies", "roadWins", "shootoutLosses", "shootoutWins",  "shutouts", "teamId", "teamName", "ties", "triCode", "wins", "total") #rename columns
  
  return(teamDF)
}
teamsT <- teamTotals() #test that function works
print(kable(head(teamsT)))  #view first 6 rows

#function that returns numeric team ID when supplied with team name
getID <- function(teamName){
  teamData <- data.frame()
  teamID <- NULL
  #searches for team name in both columns and returns numeric id, as asks for valid input
  if(teamName %in% franchiseData$teamCommonName){
    teamData <- franchiseData %>% filter(franchiseData$teamCommonName==teamName)
    teamID <- teamData$id
  } else if(teamName %in% franchiseData$teamPlaceName){
    teamData <- franchiseData %>% filter(franchiseData$teamPlaceName==teamName)
    teamID <- teamData$id
  } else {print("Please enter valid team name.")}
  return(teamID)
}

seasonRecords <- function(ID){
  url <- ""
  #creates url basted on input ID
  if(is.numeric(ID)){
    url <- paste0("-season-records?cayenneExp=franchiseId=", ID)
  } else if(is.character(ID)){
    ID = getID(ID)
    url <- paste0("-season-records?cayenneExp=franchiseId=", ID)
  }
  seasonAPI <- GET(getUrl(url)) #get url & data
  seasonText <- content(seasonAPI, "text") #convert to JSON text form
  seasonList <- fromJSON(seasonText, flatten=TRUE) #convert form
  seasonDF <- as.data.frame(seasonList) #convert to df, adds .data to col names
  colnames(seasonDF) <- c("id", "fewestGoals", "fewestGoalsAgainst", "fewestGoalsAgainstSeasons", "fewestGoalsSeasons", "fewestLosses", "fewestLossesSeasons", "fewestPoints", "fewestPointsSeasons", "fewestTies", "fewestTiesSeasons", "fewestWins", "fewestWinsSeasons", "franchiseId", "franchiseName", "homeLossStreak", "homeLossStreakDates", "homePointStreak", "homePointStreakDates", "homeWinStreak", "homeWinStreakDates", "homeWinlessStreak", "homeWinlessStreakDates", "lossStreak", "lossStreakDates", "mostGameGoals", "mostGameGoalsDates", "mostGoals", "mostGoalsAgainst", "mostGoalsAgainstSeasons", "mostGoalsSeasons", "mostLosses", "mostLossesSeasons", "mostPenaltyMinutes", "mostPenaltyMinutesSeasons", "mostPoints", "mostPointsSeasons", "mostShutouts", "mostShutoutsSeasons", "mostTies", "mostTiesSeasons", "mostWins", "mostWinsSeasons", "pointStreak", "pointStreakDates", "roadLossStreak", "roadLossStreakDates", "roadPointStreak", "roadPointStreakDates", "roadWinStreak", "roadWinStreakDates", "roadWinlessStreak", "roadWinlessStreakDates", "winStreak", "winStreakDates", "winlessStreak", "winlessStreakDates") #change df col names

  return(seasonDF)
}
seasonRecord <- seasonRecords("Carolina") #test that function works
print(kable(head(seasonRecord)))

goalieRecords <- function(ID){
  url <- ""
  #creates url basted on input ID
  if(is.numeric(ID)){
    url <- paste0("-goalie-records?cayenneExp=franchiseId=", ID)
  } else if(is.character(ID)){
    ID = getID(ID)
    url <- paste0("-goalie-records?cayenneExp=franchiseId=", ID)
  }
  goalieAPI <- GET(getUrl(url)) #get url & data
  goalieText <- content(goalieAPI, "text") #convert to JSON text form
  goalieList <- fromJSON(goalieText, flatten=TRUE) #convert form
  goalieDF <- as.data.frame(goalieList) #convert to df, adds .data to col names
  colnames(goalieDF) <- c("id", "activePlayer", "firstName", "franchiseId", "franchiseName", "gameTypeId", "gamesPlayed", "lastName", "losses", "mostGoalsAgainstDates", "mostGoalsAgainstOneGame", "mostSavesDates", "mostSavesOneGame", "mostShotsAgainstDates", "mostShotsAgainstOneGame", "mostShutoutsOneSeason", "mostShutoutsSeasonIds", "mostWinsOneSeason", "mostWinsSeasonIds", "overtimeLosses", "playerId", "positionCode", "rookieGamesPlayed", "rookieShutouts", "rookieWins", "seasons", "shutouts", "ties", "dwins", "total") #change df col names

  return(goalieDF)
}
goalieRecord <- goalieRecords("Carolina") #test that function works
print(kable(head(goalieRecord)))

skaterRecords <- function(ID){
  url <- ""
  #creates url basted on input ID
  if(is.numeric(ID)){
    url <- paste0("-skater-records?cayenneExp=franchiseId=", ID)
  } else if(is.character(ID)){
    ID = getID(ID)
    url <- paste0("-skater-records?cayenneExp=franchiseId=", ID)
  }
  skaterAPI <- GET(getUrl(url)) #get url & data
  skaterText <- content(skaterAPI, "text") #convert to JSON text form
  skaterList <- fromJSON(skaterText, flatten=TRUE) #convert form
  skaterDF <- as.data.frame(skaterList) #convert to df, adds .data to col names

  colnames(skaterDF) <- c("id", "activePlayer", "assists", "firstName", "franchiseId", "franchiseName", "gameTypeId", "gamesPlayed", "goals", "lastName", "mostAssistsGameDates", "mostAssistsOneGame", "mostAssistsOneSeason", "mostAssistsSeasonIds", "mostGoalsGameDates", "mostGoalsOneGame", "mostGoalsOneSeason", "mostGoalsSeasonIds", "mostPenaltyMinutesOneSeason", "mostPenaltyMinutesSeasonIds", "mostPointsGameDates", "mostPointsOneGame", "mostPointsOneSeason", "mostPointsSeasonIds", "penaltyMinutes", "playerId", "points", "positionCode", "rookiePoints", "seasons", "total") #change df col names

  return(skaterDF)
}
skaterRecord <- skaterRecords(9) #test that function works
print(kable(head(skaterRecord)))
```
  
### Stats API  
The user is required to input a team ID or name and modifier to access the desired data. This string input format isn't ideal, but is corrected in the wrapper function.  
Input guide for modifiers, copied from API website:  
  
  - ?expand=team.roster Shows roster of active players for the specified team  
  - ?expand=person.names Same as above, but gives less info  
  - ?expand=team.schedule.next Returns details of the upcoming game for a team  
  - ?expand=team.schedule.previous Same as above but for the last game played  
  - ?expand=team.stats Returns the teams stats for the season  
  - ?expand=team.roster&season=20142015 Adding the season identifier shows the roster for that season  
  - ?teamId=4,5,29 Can string team id together to get multiple teams  
```{r statsAPI, message=FALSE}
#get data from stats API
teams <- function(ID, input){
  url <- ""
  #creates url basted on input ID
  if(is.numeric(ID)){
    url <- paste0("https://statsapi.web.nhl.com/api/v1/teams/", ID, "/", input)
  } else if(is.character(ID)){
    ID = getID(ID)
    url <- paste0("https://statsapi.web.nhl.com/api/v1/teams/", ID, "/", input)
  }
  teamsAPI <- GET(url)
  teamsText <- content(teamsAPI, "text") #convert to JSON text form
  teamsList <- fromJSON(teamsText, flatten=TRUE) #convert form
  teamsDF <- as.data.frame(teamsList) #convert to df

  return(teamsDF)
}
teamsRecord <- teams(12,"?expand=person.names") #test that function works
print(kable(teamsRecord))

teamsRecord2 <- teams(54,"?expand=team.schedule.next") #test that function works
print(kable(teamsRecord2))

#API returns the same team information table across the different modifiers with nested data frame(s) that contain the modifier data. On the discussion board, the instructor noted we can print the output as is and that we shouldn't worry about the last modifier, so I omitted modifier 13 (?stats=statsSingleSeasonPlayoffs).
```

## Wrapper Function
This wrapper function allows the user to access any of the API endpoints and modifiers included in the above code. The user must input the endpoint and team ID or name, where team input is applicable. The endpoints have been coded by numerbers, making user input more friendly.  
  
Endpoint input guide, number 1-5 are from the Records API and numbers 6-12 are from the Stats API:  
  - 1: franchise information for all teams  
  - 2: team totals for all teams  
  - 3: season records for a team (must specify team ID or name)  
  - 4: goalie records for a team (must specify team ID or name)  
  - 5: skater records for a team (must specify team ID or name)  
  - 6: team roster of active players for a team (must specify team ID or name)  
  - 7: active player names for a team (must specify team ID or name)  
  - 8: upcoming game for a team (must specify team ID or name)  
  - 9: last game played for a team (must specify team ID or name)  
  - 10: season stats for a team (must specify team ID or name)  
  - 11: roster by season for a team (must specify team ID or name)  
  - 12: franchise information for multiple teams (must specify team IDs or names)  
  
  
```{r wrapperFcn, message=FALSE}
#function to access endpoints of stats API
accessAPI <- function(endpoint, team=NULL){
  outputDF <- NULL #create output object
  
  #check if team name is character or valid number, get numeric ID if character
  if(is.null(team)){
  } else if(is.character(team)){
    team = getID(team)
  } else if(!is.numeric(team)){
    print("Please enter valid team name or ID.")
  }
  
  #check if endpoint is valid, get url based on endpoint input
  if(endpoint==1){
    outputDF <- franchise()
  } else if(endpoint==2){
    outputDF <- teamTotals()
  } else if(endpoint==3){
    outputDF <- seasonRecords(team)
  } else if(endpoint==4){
    outputDF <- goalieRecords(team)
  } else if(endpoint==5){
    outputDF <- skaterRecords(team)
  } else if(endpoint==6){
    outputDF <- teams(team, "?expand=team.roster")
  } else if(endpoint==7){
    outputDF <- teams(team, "?expand=person.names")
  } else if(endpoint==8){
    outputDF <- teams(team, "?expand=team.schedule.next")
  } else if(endpoint==9){
    outputDF <- teams(team, "?expand=team.schedule.previous")
  } else if(endpoint==10){
    outputDF <- teams(team, "?expand=team.stats")
  } else if(endpoint==11){
    outputDF <- teams(team, "?expand=team.roster&season=20142015")
  } else if(endpoint==12){
    outputDF <- teams(team, "?teamId=4,5,29")
  } else {
    print("Please enter a valid number endpoint or modifier between 1 and 13.")
  }
return(outputDF)
}
#test code, all work but omitted to reduce knit document size
#franchise1 <- accessAPI(1)
#print(kable(head(franchise1)))

#goalie1<- accessAPI(4, "Carolina")
#print(kable(head(goalie1)))

#roster1 <- accessAPI(6, team=12)
#print(kable(roster1))
```
## Exploratory Analysis
A basic exploratory data analysis to show what can be done with data returned from the created functions.  

### A Join
Combining data sets is common in data analysis. Here I simply added row from two different endpoints together to view what columns they have in common.  
```{r join, message=FALSE}
#join goalie records and skater records for Boston Bruins
goaliesB <- accessAPI(4, team="Boston") #get data
skatersB <- accessAPI(5, team="Boston") #get data
playersB <- bind_rows(goaliesB, skatersB)
combo <- bind_rows(head(playersB),tail(playersB)) #join first 6 and low 4 rows because viewing entire data set is too much for the knit document to create
print(kable(combo))
```

### New Variables
New variables are added to the data sets below to demonstrate how the data sets can be manipulated to give useful results. These examples provide the ratio of a goalie's goals against and team losses, and what a team's mascot's name is or was.  
```{r newVariables, message=FALSE}
#add ratio of most goals against versus saves for the Carolina Hurricanse
goaliesC <- accessAPI(4, "Carolina") #get data
goaliesC <- goaliesC %>% mutate(ratio=mostGoalsAgainstOneGame/mostSavesOneGame) #add & compute ratio
print(kable(head(goaliesC)))

#add column to Atlanta Thrashers for the mascot name
rosterA <- accessAPI(6, team=11) #get data
rosterA <- rosterA %>% mutate(mascot="Thrash") #add mascot
print(kable(rosterA))
```

### Contingency Tables
These tables provide count summaries of selected variables in the data sets. The code shows the counts of total games for all franchises by game type, a breakdown of franchises that are active or inactive, and how many games were played by goalies of the Boston Bruins.  
```{r contingencyTables, message=FALSE}
#view game type counts from team totals
#1 is pre-season, 2 is regular season, 3 is post-season")
teamTots <- accessAPI(2) #get data
table(kable(teamTots$gameTypeId))

#view active(1) vs inactive(0) franchises from team totals
table(kable(teamTots$activeFranchise))

#view how many goalies played in the same number of games for Boston Bruins
table(kable(goaliesB$gamesPlayed))
```

### Numeric Summaries
These numeric summaries are calculated from the data sets. The first code returns a 5-number summary of the total teams records, the second code provides a correlation between losses and games playbe by Carolines Hurricanes goalies, and the third code provides average wins and losses by franchise ID.  
```{r numSummeries, message=FALSE}
#summary of wins for all franchises
print(summary(teamTots$wins))

#view correlation between losses and games played by Carolina Hurricanes goalies
print(cor(goaliesC$losses, goaliesC$gamesPlayed))

#view average wins and losses by franchise id
print(kable(teamTots %>% group_by(teamTots$franchiseId) %>% summarise(avgWins=mean(wins), avgLosses=mean(losses))))
```

### Plots
These plots demonstrate visuals created from data pulled using the above functions. The bar plot shows that categorical data can be plotted to view number of active and inactive franchises in the NFL. The histogram plot shows how many times each team has changed their name since their establishment. The histogram plot is a density plot of wins for all franchise teams. The scatter plot shows how many games were played by goalies of the Carolina Hurricanes against the number of losses each goalie experienced. The box plot is a bit more interesting and depicts the number of games played by goalies of the Carolina Hurricanes, Boston Bruins, and New York Rangers.  
```{r plots, message=FALSE}
#many plots but won't add labels because some categorical data is coded numerically, but information is included where possible

#bar plot of franchise status
g <- ggplot(teamTots, aes(x=activeFranchise)) #create base plotting object
g + geom_bar(aes(fill=as.factor(activeFranchise))) + labs(x="inactive      Franchise Status      active", title="Number of Active and Inactive NHL Franchises") + scale_x_discrete(labels=c("inactive", "active")) + scale_fill_discrete(name="Status", labels=c("Inactive","Active")) #add layers

#histogram plot of name changes per franchise id
i <- ggplot(franchiseData, aes(firstSeasonId)) #create base plotting object
i + geom_histogram(aes(fill=as.factor(firstSeasonId))) + labs(x="First Season ID", title="Name Changes of NFL Teams by First Season ID") + scale_fill_discrete(name="First Season ID") #add layers

#histogram density plot of all franchise wins
h <- ggplot(teamTots, aes(x=wins)) #create base plotting object
h + geom_histogram(aes(y=..density.., fill=as.factor(franchiseId))) + labs(x="Number of Wins", title="Density Plot of All Franchise Wins by Franchise ID") + scale_fill_discrete(name="Franchise ID") + geom_density() #add layers

#scatter plot of Carolina goalie games player versus number of losses
k <- ggplot(goaliesC, aes(x=gamesPlayed, y=losses)) #create base plotting object
k + geom_point(color="purple") + labs(x="Number of Games Played", y="Number of Losses", title="Carolina Hurricane Goalie Games Played versus Losses") + geom_smooth(size=0.12) #add layers

#box plot of Carolina, Boston, NY goalies games played
goaliesR <- accessAPI(4, "Rangers") #get data
goalies2 <- bind_rows(goaliesC, goaliesB, goaliesR)
j <- ggplot(goalies2, aes(x=franchiseName, y=gamesPlayed)) #create base plotting object
j + geom_boxplot() + labs(x="Franchise Name", y="Games Played", title="Games Played by Goalies from Carolina Hurricanes, Boston Bruins, and New York Rangers") #add layers
```
