---
title: "Project 1 - Karen Lopez"
date: "September 18, 2020"
output:
  rmarkdown::github_document:
    toc: true
knit: (function(input, ...){
  rmarkdown::render(
    input,
    output_file="README.md",
    envir=globalenv()
  )})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Reading and Summarizing Data from the NHL's API
This vignette was compiled to provide detailed instructions on reading and summarizing data from the National Hockey League's, or NHL's, API using R code.

## Required Packages
First, install and read in the necessary packages, as shown below. Packages only need to be installed once.  
```{r packages, message=FALSE}
#install.packages("tidyverse")
#library(tidyverse)
```

## Contact and Get Data
This code contains functions that will contact the [NHL records API](https://gitlab.com/dword4/nhlapi/-/blob/master/records-api.md) and [NHL stat API](https://gitlab.com/dword4/nhlapi/-/blob/master/stats-api.md), and return parsed, well-formatted data.  
```{r getData}
franchise <- function(){}

teamTotals <- function(){}

seasonRecords <- function(){}

goalieRecords <- function(){}

skaterRecords <- function(){}


teams <- function(){}
```

## Wrapper Function

```{r wrapperFcn}
allAccess <- function(){}
```

## Exploratory Analysis

```{r exploreAnalysis}
#do a join

#create 2 new variables

#create some contingency tables

#create numerical summaries

#create 5 plots

```
